---
title: "R Notebook"
output: html_notebook
---

```{r setup}
library(dplyr)
library(tidyverse)
library(factoextra)
library(densityClust)
library(stats)
library(sf)
library(ggpubr)
library(grid)
library(vegan)
library(phytools)
library(picante)
rm(list = ls())
```

```{r}
col_loc = read_csv('./bigquery_export__columbidae_birdlife_urban_locations.csv')
head(col_loc)
```

```{r}
col_ut = read_csv('./bigquery_export__columbidae_birdlife_urban_tolerance.csv')
head(col_ut)
```

```{r}
avonet_birdlife = read_csv('./avonet_birdlife.csv')
head(avonet_birdlife)
```

Check we have traits for all species in our dataset
```{r}
distinct(col_loc[!col_loc$species_name %in% avonet_birdlife$Species1,])
distinct(col_ut[!col_ut$species_name %in% avonet_birdlife$Species1,])
```



Bregman et al. (2015): https://esajournals.onlinelibrary.wiley.com/doi/full/10.1890/14-1731.1
Trisos et al. (2014)
locomotory (tail, wing, and tarsus length) 
trophic (beak length, width, and depth) traits

```{r}
col_traits = avonet_birdlife[avonet_birdlife$Species1 %in% col_loc$species_name,c('Species1', 'Beak.Length_Culmen', 'Beak.Width', 'Beak.Depth', 'Tail.Length', 'Hand-Wing.Index', 'Tarsus.Length', 'Habitat', 'Habitat.Density', 'Trophic.Niche', 'Primary.Lifestyle', 'Range.Size', 'Mass')]
names(col_traits) <- c('species_name', 'beak_length', 'beak_width', 'beak_depth', 'tail_length', 'wing_index', 'tarsus_length', 'habitat', 'habitat_density', 'trophic_niche', 'primary_lifestyle', 'range_size', 'mass')
col_traits
```

```{r}
pigeons = left_join(col_ut[,c('species_name', 'urban_tolerance', 'region_count')], col_traits)
pigeons
```

```{r}
pigeons5 = pigeons[pigeons$region_count >= 5,]
```

```{r}
shapiro.test(pigeons$urban_tolerance) #p <= 0.05   -> Significantly different from normal
```

----------------------------
Trophic niches
----------------------------
```{r}
figure1 = ggarrange(ncol = 2, nrow = 1,
  ggplot(pigeons, aes(x = trophic_niche, y = urban_tolerance)) + 
    geom_boxplot() + labs(title = "All species") + rremove("ylab") + rremove("xlab"),
  ggplot(pigeons5, aes(x = trophic_niche, y = urban_tolerance)) + 
    geom_boxplot() + labs(title = "Min. 5 regional pools") + rremove("ylab") + rremove("xlab")
)

annotate_figure(figure1, left = textGrob("Urban tolerance", rot = 90, vjust = 1, gp = gpar(cex = 1.3)),
                    bottom = textGrob("Trophic niche"))
```
```{r}
kruskal.test(pigeons$urban_tolerance~pigeons$trophic_niche)
pairwise.wilcox.test(pigeons$urban_tolerance, pigeons$trophic_niche)
```

# Urban tolerance was significantly different between different trophic niches (Chi-squared=19.31, df=2, p<0.001), with urban tolerance significantly 
# lower in Frugivores compared to Granivores.

```{r}
kruskal.test(pigeons5$urban_tolerance~pigeons5$trophic_niche)
pairwise.wilcox.test(pigeons5$urban_tolerance, pigeons5$trophic_niche)
```

----------------
Habitat
----------------
```{r}
figure2 = ggarrange(ncol = 2, nrow = 1,
  ggplot(pigeons, aes(x = habitat, y = urban_tolerance)) + 
    geom_boxplot() + labs(title = "All species") + rremove("ylab") + rremove("xlab") + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)),
  ggplot(pigeons5, aes(x = habitat, y = urban_tolerance)) + 
    geom_boxplot() + labs(title = "Min. 5 regional pools") + rremove("ylab") + rremove("xlab") + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
)

annotate_figure(figure2, left = textGrob("Urban tolerance", rot = 90, vjust = 1, gp = gpar(cex = 1.3)),
                    bottom = textGrob("Habitat"))
```

```{r}
kruskal.test(pigeons$urban_tolerance~pigeons$habitat)
pairwise.wilcox.test(pigeons$urban_tolerance, pigeons$habitat)
```

# Urban tolerance was significantly different between different trophic niches (Chi-squared=42.834, df=6, p<0.001)

```{r}
kruskal.test(pigeons5$urban_tolerance~pigeons5$habitat)
pairwise.wilcox.test(pigeons5$urban_tolerance, pigeons5$habitat)
```

```{r}
figure3 = ggarrange(ncol = 2, nrow = 1,
  ggplot(pigeons, aes(x = as.factor(habitat_density), y = urban_tolerance)) + 
    geom_boxplot() + labs(title = "All species") + rremove("ylab") + rremove("xlab") + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)),
  ggplot(pigeons5, aes(x = as.factor(habitat_density), y = urban_tolerance)) + 
    geom_boxplot() + labs(title = "Min. 5 regional pools") + rremove("ylab") + rremove("xlab") + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
)

annotate_figure(figure3, left = textGrob("Urban tolerance", rot = 90, vjust = 1, gp = gpar(cex = 1.3)),
                    bottom = textGrob("Habitat Density"))
```
```{r}
kruskal.test(pigeons$urban_tolerance~pigeons$habitat_density)
pairwise.wilcox.test(pigeons$urban_tolerance, pigeons$habitat_density)
```


----------------------------
Lifestyle
----------------------------
```{r}
figure4 = ggarrange(ncol = 2, nrow = 1,
  ggplot(pigeons, aes(x = primary_lifestyle, y = urban_tolerance)) + 
    geom_boxplot() + labs(title = "All species") + rremove("ylab") + rremove("xlab"),
  ggplot(pigeons5, aes(x = primary_lifestyle, y = urban_tolerance)) + 
    geom_boxplot() + labs(title = "Min. 5 regional pools") + rremove("ylab") + rremove("xlab")
)

annotate_figure(figure4, left = textGrob("Urban tolerance", rot = 90, vjust = 1, gp = gpar(cex = 1.3)),
                    bottom = textGrob("Primary Lifestyle"))
```

```{r}
kruskal.test(pigeons$urban_tolerance~pigeons$primary_lifestyle)
pairwise.wilcox.test(pigeons$urban_tolerance, pigeons$primary_lifestyle)
```

# Urban tolerance was significantly different between different trophic niches (Chi-squared=7.23, df=2, p = 0.03), with urban tolerance significantly 
# lower in Insessorial (perching) compared to Terrestrial and Generalist

```{r}
kruskal.test(pigeons5$urban_tolerance~pigeons5$primary_lifestyle)
pairwise.wilcox.test(pigeons5$urban_tolerance, pigeons5$primary_lifestyle)
```

--------------------------
Create at locomotion trait
--------------------------
```{r}
loco_pca = rda(pigeons[,c('tarsus_length', 'wing_index', 'tail_length')])
pigeons$loco_pc1 = loco_pca$Ybar[,1]
pigeons$loco_pc2 = loco_pca$Ybar[,2]
pigeons$loco_pc3 = loco_pca$Ybar[,3]
summary(loco_pca)
```

```{r}
ggarrange(ncol = 4, nrow = 3,
  ggplot(pigeons, aes(x = loco_pc1, y = tarsus_length)) + geom_point() + labs(subtitle = cor(pigeons$loco_pc1,pigeons$tarsus_length)),
  ggplot(pigeons, aes(x = loco_pc1, y = wing_index)) + geom_point() + labs(subtitle = cor(pigeons$loco_pc1,pigeons$wing_index)),
  ggplot(pigeons, aes(x = loco_pc1, y = tail_length)) + geom_point() + labs(subtitle = cor(pigeons$loco_pc1,pigeons$tail_length)),
  ggplot(pigeons, aes(x = loco_pc1, y = mass)) + geom_point() + labs(subtitle = cor(pigeons$loco_pc1,pigeons$mass)),
  ggplot(pigeons, aes(x = loco_pc2, y = tarsus_length)) + geom_point() + labs(subtitle = cor(pigeons$loco_pc2,pigeons$tarsus_length)),
  ggplot(pigeons, aes(x = loco_pc2, y = wing_index)) + geom_point() + labs(subtitle = cor(pigeons$loco_pc2,pigeons$wing_index)),
  ggplot(pigeons, aes(x = loco_pc2, y = tail_length)) + geom_point() + labs(subtitle = cor(pigeons$loco_pc2,pigeons$tail_length)),
  ggplot(pigeons, aes(x = loco_pc2, y = mass)) + geom_point() + labs(subtitle = cor(pigeons$loco_pc2,pigeons$mass)),
  ggplot(pigeons, aes(x = loco_pc3, y = tarsus_length)) + geom_point() + labs(subtitle = cor(pigeons$loco_pc3,pigeons$tarsus_length)),
  ggplot(pigeons, aes(x = loco_pc3, y = wing_index)) + geom_point() + labs(subtitle = cor(pigeons$loco_pc3,pigeons$wing_index)),
  ggplot(pigeons, aes(x = loco_pc3, y = tail_length)) + geom_point() + labs(subtitle = cor(pigeons$loco_pc3,pigeons$tail_length)),
  ggplot(pigeons, aes(x = loco_pc3, y = mass)) + geom_point() + labs(subtitle = cor(pigeons$loco_pc3,pigeons$mass))
)
```

--------------------------
Create at trophic trait
--------------------------
```{r}
trophic_pca = rda(pigeons[,c('beak_length', 'beak_width', 'beak_depth')])
pigeons$trophic_pc1 = trophic_pca$Ybar[,1]
pigeons$trophic_pc2 = trophic_pca$Ybar[,2]
pigeons$trophic_pc3 = trophic_pca$Ybar[,3]
summary(trophic_pca)
```

```{r}
ggarrange(ncol = 4, nrow = 3,
  ggplot(pigeons, aes(x = trophic_pc1, y = beak_length)) + geom_point() + labs(subtitle = cor(pigeons$trophic_pc1,pigeons$beak_length)),
  ggplot(pigeons, aes(x = trophic_pc1, y = beak_width)) + geom_point() + labs(subtitle = cor(pigeons$trophic_pc1,pigeons$beak_width)),
  ggplot(pigeons, aes(x = trophic_pc1, y = beak_depth)) + geom_point() + labs(subtitle = cor(pigeons$trophic_pc1,pigeons$beak_depth)),
  ggplot(pigeons, aes(x = trophic_pc1, y = mass)) + geom_point() + labs(subtitle = cor(pigeons$trophic_pc1,pigeons$mass)),
  ggplot(pigeons, aes(x = trophic_pc2, y = beak_length)) + geom_point() + labs(subtitle = cor(pigeons$trophic_pc2,pigeons$beak_length)),
  ggplot(pigeons, aes(x = trophic_pc2, y = beak_width)) + geom_point() + labs(subtitle = cor(pigeons$trophic_pc2,pigeons$beak_width)),
  ggplot(pigeons, aes(x = trophic_pc2, y = beak_depth)) + geom_point() + labs(subtitle = cor(pigeons$trophic_pc2,pigeons$beak_depth)),
  ggplot(pigeons, aes(x = trophic_pc2, y = mass)) + geom_point() + labs(subtitle = cor(pigeons$trophic_pc2,pigeons$mass)),
  ggplot(pigeons, aes(x = trophic_pc3, y = beak_length)) + geom_point() + labs(subtitle = cor(pigeons$trophic_pc3,pigeons$beak_length)),
  ggplot(pigeons, aes(x = trophic_pc3, y = beak_width)) + geom_point() + labs(subtitle = cor(pigeons$trophic_pc3,pigeons$beak_width)),
  ggplot(pigeons, aes(x = trophic_pc3, y = beak_depth)) + geom_point() + labs(subtitle = cor(pigeons$trophic_pc3,pigeons$beak_depth)),
  ggplot(pigeons, aes(x = trophic_pc3, y = mass)) + geom_point() + labs(subtitle = cor(pigeons$trophic_pc3,pigeons$mass))
)
```

--------------------------------------------------
Look at traits distribution using picante
---------------------------------------------------

```{r}
mapping = read_csv('./bigquery_export__columbidae_taxonomic_mapping.csv')
```

```{r}
pigeons_jetz = left_join(pigeons, mapping)
pigeons_jetz = pigeons_jetz %>% group_by(jetz_scientific_name) %>% summarise(locomotion = mean(loco_pc1), trophic = mean(trophic_pc1))
picante_traits = pigeons_jetz[,c('jetz_scientific_name', 'locomotion', 'trophic')]
picante_traits = tibble::column_to_rownames(picante_traits, var='jetz_scientific_name')
picante_traits = as.matrix(picante_traits)
```

```{r}
tree <- read.tree("./phylogeny__stage2_hackett_mcc_no_neg.tre")
```

```{r}
tree_pruned_to_columbidae <- ladderize(drop.tip(tree, setdiff(tree$tip.label, rownames(picante_traits))))
```

```{r}
phydist <- cophenetic(tree_pruned_to_columbidae)
```

```{r}
par(mfrow=c(1,2))
plot(tree_pruned_to_columbidae, show.tip.label=FALSE, main='locomotion')
tiplabels(pch=22, col=picante_traits[,1], bg=picante_traits[,1], cex=1.5)

plot(tree_pruned_to_columbidae, show.tip.label=FALSE, main='trophic')
tiplabels(pch=22, col=picante_traits[,2], bg=picante_traits[,0], cex=1.5)
```
```{R}
ggplot(pigeons, aes(x = loco_pc1, y = trophic_pc1, color = urban_tolerance)) + geom_point() + scale_color_gradientn(colours = terrain.colors(7))
```

```{r}
ggplot(pigeons, aes(x = loco_pc1, y = trophic_pc1, color = trophic_niche)) + geom_point()
```
```{r}
summary(glm(data = pigeons, formula = as.factor(trophic_niche) ~ loco_pc1 * trophic_pc1, family = binomial))
```

```{r}
ggplot(pigeons, aes(x = loco_pc1, y = trophic_pc1, color = primary_lifestyle)) + geom_point()
```

```{r}
summary(glm(data = pigeons[pigeons$primary_lifestyle != 'Generalist',], formula = as.factor(primary_lifestyle) ~ loco_pc1 * trophic_pc1, family = binomial))
```

```{r}
ggplot(pigeons, aes(x = loco_pc1, y = trophic_pc1, color = habitat)) + geom_point()
```

```{r}
summary(glm(data = pigeons, formula = as.factor(habitat) ~ loco_pc1 * trophic_pc1, family = binomial))
```
