---
title: "Presence Clustering by regional pool"
output: html_notebook
---

Can we cluster species together based on the regional pools they co-occur in?
And then work out whether there are common species present/absent from those clusters?

```{r setup}
library(dplyr)
library(tidyverse)
library(factoextra)
library(densityClust)
library(stats)
library(sf)
rm(list = ls())

col_loc <- read_csv('./bigquery_export__columbidae_jetz_urban_locations.csv')
```


```{r}
head(col_loc)
```
## Create matrix of species by city, with 1 as in regional pool, and 0 as not in regional pool
```{r}
regional_pool_data = distinct(col_loc[,c('city_name', 'jetz_scientific_name')])
regional_pool_data$city_name = str_replace_all(regional_pool_data$city_name, '[ ,-]+', '_')
regional_pool_data$present = 1
species_in_pools = pivot_wider(regional_pool_data, names_from = city_name, values_from = "present", values_fill = list(present = 0))
species_in_pools = tibble::column_to_rownames(species_in_pools, var='jetz_scientific_name')
species_in_pools
```
## What is the optimium number of clusters for this matrix?
```{r}
fviz_nbclust(species_in_pools, kmeans, method = "gap_stat", k.max = 25)
```
## Let's create 10 clusters then
```{r}
clusters = kmeans(species_in_pools, 10)

clustered_species = as.data.frame(clusters$cluster)
clustered_species$jetz_scientific_name = rownames(clustered_species)
rownames(clustered_species) <- NULL
names(clustered_species) <- c('cluster', 'jetz_scientific_name')
head(clustered_species)
```
### How many clusters are in London?
```{r}
cities_to_clusters = left_join(col_loc, clustered_species)
cities_to_clusters[cities_to_clusters$city_name == 'London',]
```
### What cities does cluster 1 cover?
```{r}
unique(cities_to_clusters[cities_to_clusters$cluster == 1,c('city_name')])
```
### What cities does cluster 7 cover?
```{r}
unique(cities_to_clusters[cities_to_clusters$cluster == 10,c('city_name')])
```
## Let's look at urban tolerance in each species
```{r}
col_ut <- read_csv('./bigquery_export__columbidae_jetz_urban_tolerance.csv')
```
```{r}
clustered_species_ut = left_join(col_ut, clustered_species)
clustered_species_ut
```

## Do we have species with differing amounts of urban tolerance?  We would expect so if dominant species populate cities over others
```{r, fig.width=12}
library(ggrepel)
ggplot(clustered_species_ut, aes(x = 1, y = urban_tolerance)) + 
  geom_text_repel(aes(label = jetz_scientific_name)) + facet_wrap(~ cluster, ncol = 5)
```
## Let's look at what geographical area these clusters cover
```{r}
clustered_species_loc = left_join(col_loc, clustered_species)
head(clustered_species_loc)
```

```{r}
world_distributions <- read_sf('./urban_centres/urban_centres.shp')
world_cities <- distinct(world_distributions[,c('fid', 'geometry')])
names(world_cities) <- c('city_id', 'geometry')
head(world_cities)
```

```{r}
clustered_species_cities = left_join(clustered_species_loc, world_cities)
head(clustered_species_cities)
```
```{r}
cluster_7_points = clustered_species_cities[clustered_species_cities$cluster == 7, ]
cluster_7_feature_list = st_sf(cluster_7_points, sf_column_name = 'geometry')
st_as_sfc(st_bbox(cluster_7_feature_list))
```

```{r}
country_outlines <- read_sf('./world_country_outlines/world-administrative-boundaries.shp')
country_outlines <- st_transform(country_outlines, crs = st_crs(world_distributions))
```

```{r}
bounding_boxes = data.frame()
crs = st_crs(world_distributions)

for (cluster in 1:10) {
  points = clustered_species_cities[clustered_species_cities$cluster == cluster, c('geometry')]
  box = st_bbox(st_sf(points, sf_column_name = 'geometry'), crs = crs)
  bounding_boxes = rbind(bounding_boxes, data.frame(cluster = cluster, geometry = st_as_sfc(box)))
}

bounding_boxes = st_sf(bounding_boxes)
bounding_boxes
```


```{r}
ggplot() +  
  geom_sf(data = country_outlines, aes(geometry = geometry)) + 
  geom_sf(data = bounding_boxes, aes(geometry = geometry, color = as.factor(cluster)), alpha = 0.2)
ggsave('clustering_by_pool__region_by_cluster.jpg')
```

## How do traits occur within these clusters given urban tolerance?
```{r}
avonet_birdlife = read_csv('./avonet_birdlife.csv')
head(avonet_birdlife)
```
```{r}
mapping = read_csv('./bigquery_export__columbidae_taxonomic_mapping.csv')
```

```{r}
avonet_mapping = left_join(mapping, avonet_birdlife, by = c('species_name' = 'Species1'))
avonet_jetz_dups = avonet_mapping[, c('jetz_scientific_name', 'Beak.Length_Culmen', 'Beak.Length_Nares', 'Beak.Width', 'Beak.Depth', 'Tarsus.Length', 'Wing.Length', 'Kipps.Distance', 'Secondary1', 'Hand-Wing.Index', 'Tail.Length', 'Habitat.Density', 'Trophic.Niche', 'Primary.Lifestyle')]

avonet_jetz = avonet_jetz_dups %>% group_by(jetz_scientific_name) %>% summarise(
  beak_length_culmen = mean(Beak.Length_Culmen),
  beak_length_nares = mean(Beak.Length_Nares),
  beak_width = mean(Beak.Width),
  beak_depth = mean(Beak.Depth),
  tarsus_length = mean(Tarsus.Length),
  wing_length = mean(Wing.Length),
  kipps_distance = mean(Kipps.Distance),
  secondary = mean(Secondary1),
  hand_wing_index = mean(`Hand-Wing.Index`),
  tail_length = mean(Tail.Length)
)
avonet_jetz
```
```{r}
col_traits_ut = left_join(clustered_species_ut, avonet_jetz)
col_traits_ut
```

```{r}
library(scales)
col_traits_ut$beak_length_culmen_scaled = rescale(col_traits_ut$beak_length_culmen, c(0, 1))
col_traits_ut$beak_length_nares_scaled = rescale(col_traits_ut$beak_length_nares, c(0, 1))
col_traits_ut$beak_width_scaled = rescale(col_traits_ut$beak_width, c(0, 1))
col_traits_ut$beak_depth_scaled = rescale(col_traits_ut$beak_depth, c(0, 1))
```

```{r}
feeding_traits = pivot_longer(col_traits_ut[,c('jetz_scientific_name', 'urban_tolerance', 'cluster', 'beak_length_culmen_scaled', 'beak_length_nares_scaled', 'beak_width_scaled', 'beak_depth_scaled')],
             cols = c('beak_length_culmen_scaled', 'beak_length_nares_scaled', 'beak_width_scaled', 'beak_depth_scaled'), names_to = 'trait', values_to = 'value')
feeding_traits
```

```{r}
ggplot(feeding_traits, aes(y = urban_tolerance, x = as.factor(cluster))) + geom_point(aes(size = value, color = value)) + facet_wrap(~trait) + scale_color_gradientn(colours = terrain.colors(7))
ggsave('clustering_by_pool__trophic_trait.jpg')

```

```{r}
col_traits_ut$wing_length_scaled = rescale(col_traits_ut$wing_length, c(0, 1))
col_traits_ut$secondary_scaled = rescale(col_traits_ut$secondary, c(0, 1))
col_traits_ut$hand_wing_index_scaled = rescale(col_traits_ut$hand_wing_index, c(0, 1))
col_traits_ut$tail_length_scaled = rescale(col_traits_ut$tail_length, c(0, 1))
```

```{r}
locomotion_traits = pivot_longer(col_traits_ut[,c('jetz_scientific_name', 'urban_tolerance', 'cluster', 'wing_length_scaled', 'secondary_scaled', 'hand_wing_index_scaled', 'tail_length_scaled')],
             cols = c('wing_length_scaled', 'secondary_scaled', 'hand_wing_index_scaled', 'tail_length_scaled'), names_to = 'trait', values_to = 'value')
locomotion_traits
```

```{r}
ggplot(locomotion_traits, aes(y = urban_tolerance, x = as.factor(cluster))) + geom_point(aes(size = value, color = value)) + facet_wrap(~trait) + scale_color_gradientn(colours = terrain.colors(7))
ggsave('clustering_by_pool__dispersal_trait.jpg')
```


