---
title: "Maps of Columbidae"
output: html_notebook
---

# Map cities that pigeons occur in around globe

```{r setup}
library(dplyr)
library(tidyverse)
library(sf)
rm(list = ls())
```

### Constants
```{r}
PRESENT_SHAPE = 20
ABSENT_SHAPE = 4
PRESENT_COLOUR = "blue"
ABSENT_COLOUR = "red"
```

## Load required data
```{r}
col_ut <- read_csv('./bigquery_export__columbidae_jetz_urban_tolerance.csv')
```

```{r}
col_loc <- read_csv('./bigquery_export__columbidae_jetz_urban_locations.csv')
```

```{r}
country_outlines <- read_sf('./world_country_outlines/world-administrative-boundaries.shp')
country_outlines
```

```{r}
st_crs(country_outlines)
```

```{r}
ggplot() +  
  geom_sf(data = country_outlines, aes(geometry = geometry))
```

```{r}
col_ut %>% arrange(urban_tolerance)
```

```{r}
world_distributions <- read_sf('./urban_centres/urban_centres.shp')
world_cities <- distinct(world_distributions[,c('fid', 'geometry')])
names(world_cities) <- c('city_id', 'geometry')
world_cities
```

```{r}
ggplot() +  
  geom_sf(data = world_cities, aes(geometry = geometry))
```

```{r}
st_crs(world_cities)
```

```{r}
world_cities <- st_transform(world_cities, crs = st_crs(country_outlines))
```

# Plot locations cities with pigeons with zero urban tolerance
```{r}
intolerance_absence_records = col_loc[col_loc$jetz_scientific_name %in% col_ut$jetz_scientific_name[col_ut$urban_tolerance == 0],]
intolerance_absence_records
```

```{r}
absent_cities = left_join(intolerance_absence_records, world_cities)
absent_cities
```

```{r}
ggplot() +  
  geom_sf(data = country_outlines, aes(geometry = geometry)) +
  geom_sf(data = absent_cities, aes(geometry = geometry), shape = ABSENT_SHAPE, color = ABSENT_COLOUR, show.legend = "point") +
  labs(title = "Columbidae with zero urban tolerance") + 
  theme_bw()
```

# Plot locations cities with pigeons with full urban tolerance
```{r}
tolerant_presence_records = col_loc[col_loc$jetz_scientific_name %in% col_ut$jetz_scientific_name[col_ut$urban_tolerance == 1],]
tolerant_presence_records[tolerant_presence_records$jetz_scientific_name != 'Columba_livia',]
```

```{r}
present_cities = left_join(tolerant_presence_records[tolerant_presence_records$jetz_scientific_name != 'Columba_livia',], world_cities)
present_cities
```

```{r}
ggplot() +  
  geom_sf(data = country_outlines, aes(geometry = geometry)) +
  geom_sf(data = present_cities, aes(geometry = geometry), shape = PRESENT_SHAPE, color = PRESENT_COLOUR, show.legend = "point") +
  labs(title = "Columbidae with full urban tolerance", subtitle = "Except Columba livia") + 
  theme_bw()
```

# Plot locations of most common pigeons

## Create species title for presence records
```{r}
col_ut$title = paste(str_replace(col_ut$jetz_scientific_name, '_', ' '), ' (', col_ut$urban_tolerance, ')', sep = '')
head(col_ut$title)
```

## Constants
```{r}
TOP_X = 16
```

```{r}
most_common = col_ut[col_ut$jetz_scientific_name != 'Columba_livia',] %>% arrange(desc(city_count)) %>% head(TOP_X)
most_common
```

```{r}
most_common_locations = col_loc[col_loc$jetz_scientific_name %in% most_common$jetz_scientific_name,]
most_common_locations
```

```{r}
most_common_locations[most_common_locations$jetz_scientific_name == 'Columba_oenas',]
```

```{r}
most_common_cities = left_join(most_common_locations, world_cities)
most_common_cities = left_join(most_common_cities, col_ut[,c('jetz_scientific_name', 'title')])
most_common_cities
```

```{r, fig.width = 12, fig.height = 20}
ggplot() +  
  geom_sf(data = country_outlines, aes(geometry = geometry)) +
  geom_sf(data = most_common_cities, aes(geometry = geometry, shape = present_in_city, color = present_in_city), show.legend = "point") +
  facet_wrap(~ title, ncol = 2) +
  scale_colour_manual(name = "Presence",
                     limits = c(F, T),
                      labels = c("Regional pool", "City"),
                      values = c(ABSENT_COLOUR, PRESENT_COLOUR)) +   
  scale_shape_manual(name = "Presence",
                     limits = c(F, T),
                     labels = c("Regional pool", "City"),
                     values = c(ABSENT_SHAPE, PRESENT_SHAPE)) +
  theme_bw() + 
  theme(legend.position="bottom") + 
  labs(title = paste(TOP_X, "most commonly occuring Columbidae"), subtitle = "Except Columba livia")
```


# Investigate seasonal distributions, are these related to presence?
```{r}
dist <- read_sf('./distributions/columbidae.shp')
head(dist)
```


```{r}
dist$season_factor = factor(dist$seasonal, levels = c(1, 2, 3, 4, 5), labels = c("Resident", "Breeding", "Non-breeding", "Passage", "Uncertain"))
```

```{r}
mapping <- read_csv('./bigquery_export__columbidae_taxonomic_mapping.csv')
dist = left_join(dist, mapping, by = c('binomial' = 'species_name'))
head(dist)
```



```{r, fig.width = 12, fig.height = 20}
plot_species = function(species_list_to_city_id, plot_title) {
  cities = left_join(species_list_to_city_id, col_loc[,c('jetz_scientific_name', 'present_in_city', 'city_id')])
  cities = left_join(cities, world_cities[,c('city_id', 'geometry')])
  cities = left_join(cities, col_ut[,c('jetz_scientific_name', 'title')])
  
  dists = left_join(species_list_to_city_id, dist[, c('jetz_scientific_name', 'season_factor', 'geometry')])
  dists = left_join(dists, col_ut[,c('jetz_scientific_name', 'title')])
  
  ggplot() +  
    geom_sf(data = country_outlines, aes(geometry = geometry)) +
    geom_sf(data = dists, aes(geometry = geometry, fill = season_factor), alpha = 0.2) +
    geom_sf(data = cities, aes(geometry = geometry, shape = present_in_city, color = present_in_city)) +
    facet_wrap(~ title, ncol = 2) +
    scale_colour_manual(name = "Presence",
                       limits = c(F, T),
                        labels = c("Regional pool", "City"),
                        values = c(ABSENT_COLOUR, PRESENT_COLOUR)) +   
    scale_shape_manual(name = "Presence",
                       limits = c(F, T),
                       labels = c("Regional pool", "City"),
                       values = c(ABSENT_SHAPE, PRESENT_SHAPE)) +
    theme_bw() + 
    theme(legend.position="bottom") + 
    labs(title = paste(TOP_X, plot_title), subtitle = "Except Columba livia") +
    guides(fill=guide_legend(title="Distribution"))
}
```


## Most common
```{r, fig.width = 12, fig.height = 20}
plot_species(most_common[,c('jetz_scientific_name')], 'most commonly occuring Columbidae') 
ggsave("outputs/map_most_common.jpg")
```
## Least urban tolerant
```{r, fig.width = 12, fig.height = 20}
least_tolerant = col_ut[col_ut$jetz_scientific_name != 'Columba_livia',] %>% filter(region_count > 5) %>% arrange(urban_tolerance) %>% head(TOP_X)
plot_species(least_tolerant[,c('jetz_scientific_name')], 'least urban tolerant Columbidae')
ggsave("outputs/map_least_tolerant.jpg")
```


## Most urban tolerant
```{r, fig.width = 12, fig.height = 20}
most_tolerant = col_ut[col_ut$jetz_scientific_name != 'Columba_livia',] %>% filter(region_count > 5) %>% arrange(desc(urban_tolerance)) %>% head(TOP_X)
plot_species(most_tolerant[,c('jetz_scientific_name')], 'most urban tolerant Columbidae') 
ggsave("outputs/map_most_tolerant.jpg")
```