---
title: "Phylogeny of Columbidae"
output: html_notebook
---

```{r setup}
library(phytools)
library(dplyr)
library(tidyverse)
rm(list = ls())
```


Load a maximum clade credibility tree from jetz
```{r}
tree <- read.tree("./phylogeny__stage2_hackett_mcc_no_neg.tre")
```

Load urban tolerance data
```{r}
col_ut <- read_csv('./bigquery_export__columbidae_urban_tolerance.csv')
```

Check for species from data not in tree
```{r}
col_ut[!col_ut$jetz_scientific_name %in% tree$tip.label,]
```


Format data
```{r}
columbidae_input_species <- col_ut[col_ut$jetz_scientific_name %in% tree$tip.label,]
tree_pruned_to_columbidae <- ladderize(drop.tip(tree, setdiff(tree$tip.label, col_ut$jetz_scientific_name)))

columbidae_response <- columbidae_input_species$presence_ratio
names(columbidae_response) <- columbidae_input_species$jetz_scientific_name

head(columbidae_response)
```

Is the a phylogenetic signal for urban tolerance in the columbidae family
```{r}
col_ut_phylo_signal <- phylosig(tree_pruned_to_columbidae, columbidae_response, method="lambda", test=TRUE)
col_ut_phylo_signal
```

Show the results for species
```{r}
col_continuous_trait_evolution <- contMap(tree_pruned_to_columbidae, columbidae_response)

colourCount <- length(col_continuous_trait_evolution$cols)
col_continuous_trait_evolution$cols[1:colourCount] <- rev(col_continuous_trait_evolution$col[1:colourCount])
```

```{r}
plot(col_continuous_trait_evolution, outline=FALSE, fsize=0.3, lwd=1, leg.txt="Cities/Regional Pools", type = "fan")
```

----
What about species in very few regional pools?
----------------------------------------------
```{r}
ggplot(col_ut, aes(x = region_count)) + geom_bar()
```
Format data with species in more than 2 cities
```{r}
columbidae_reduced_input_species <- col_ut[col_ut$region_count > 2,]
columbidae_reduced_input_species <- columbidae_reduced_input_species[columbidae_reduced_input_species$jetz_scientific_name %in% tree$tip.label,]
tree_pruned_to_columbidae_reduced <- ladderize(drop.tip(tree, setdiff(tree$tip.label, columbidae_reduced_input_species$jetz_scientific_name)))

columbidae_reduced_response <- columbidae_reduced_input_species$presence_ratio
names(columbidae_reduced_response) <- columbidae_reduced_input_species$jetz_scientific_name

head(columbidae_reduced_response)
```

Is the a phylogenetic signal for urban tolerance in the columbidae family with species in at least 2 regional pools
```{r}
col_ut_reduced_phylo_signal <- phylosig(tree_pruned_to_columbidae_reduced, columbidae_reduced_response, method="lambda", test=TRUE)
col_ut_reduced_phylo_signal
```

Are species in low numbers of regional pool mostly present or absent?
```{r}
ggplot(col_ut[col_ut$region_count < 3,], aes(x = presence_ratio)) + geom_bar()
```
Is presence explained by number of regions?
-------------------------------------------
```{r}
summary(lm(formula = presence_ratio ~ region_count, data = col_ut))
```

What about for low numbers of regional pools?
```{r}
summary(lm(formula = presence_ratio ~ region_count, data = col_ut[col_ut$region_count < 3,]))
```

What about for higher numbers of regional pools?
```{r}
summary(lm(formula = presence_ratio ~ region_count, data = col_ut[col_ut$region_count > 2, ]))
```
