---
title: "Phylogeny of Columbidae"
output: html_notebook
---

```{r setup}
library(phytools)
library(dplyr)
library(tidyverse)
rm(list = ls())
```


Load a maximum clade credibility tree from jetz
```{r}
tree <- read.tree("./phylogeny__stage2_hackett_mcc_no_neg.tre")
```

Load urban tolerance data
```{r}
col_ut <- read_csv('./bigquery_export__columbidae_jetz_urban_tolerance.csv')
```

Check for species from data not in tree
```{r}
col_ut[!col_ut$jetz_scientific_name %in% tree$tip.label,]
```

Number of species in analysis:
```{r}
nrow(col_ut)
```

Format data
```{r}
columbidae_input_species <- col_ut[col_ut$jetz_scientific_name %in% tree$tip.label,]
tree_pruned_to_columbidae <- ladderize(drop.tip(tree, setdiff(tree$tip.label, col_ut$jetz_scientific_name)))

columbidae_response <- columbidae_input_species$urban_tolerance
names(columbidae_response) <- columbidae_input_species$jetz_scientific_name

head(columbidae_response)
```

Is the a phylogenetic signal for urban tolerance in the columbidae family
```{r}
col_ut_phylo_signal <- phylosig(tree_pruned_to_columbidae, columbidae_response, method="lambda", test=TRUE)
col_ut_phylo_signal
```

```
-
- Phylogenetic analysis containing all 160 species and produced a significant phylogenetic signal for urban tolerance (lambda: 0.16, LR(lambda=0): 2.34, p = 0.13)
-
```

Show the results for species
```{r}
col_continuous_trait_evolution <- contMap(tree_pruned_to_columbidae, columbidae_response)

colourCount <- length(col_continuous_trait_evolution$cols)
col_continuous_trait_evolution$cols[1:colourCount] <- rev(col_continuous_trait_evolution$col[1:colourCount])
```

```{r}
cte_all = plot(col_continuous_trait_evolution, outline=FALSE, fsize=0.3, lwd=1, leg.txt="Cities/Regional Pools", type = "fan")
cte_all
```
## Find the age of each species (tip branch length)
```{r}
species_ages = tree_pruned_to_columbidae$edge.length[tree_pruned_to_columbidae$edge[, 2] %in% 1:Ntip(tree_pruned_to_columbidae)]
tips <- tree_pruned_to_columbidae$edge[tree_pruned_to_columbidae$edge[, 2] %in%  1:Ntip(tree_pruned_to_columbidae), 2]
names(species_ages) = tree_pruned_to_columbidae$tip.label[tips]
species_ages
```

## What about species in very few regional pools?
```{r}
ggplot(col_ut, aes(x = region_count)) + geom_bar()
```

Format data with species in more than 2 cities
```{r}
columbidae_reduced_input_species <- col_ut[col_ut$region_count > 5,]
columbidae_reduced_input_species <- columbidae_reduced_input_species[columbidae_reduced_input_species$jetz_scientific_name %in% tree$tip.label,]
tree_pruned_to_columbidae_reduced <- ladderize(drop.tip(tree, setdiff(tree$tip.label, columbidae_reduced_input_species$jetz_scientific_name)))

columbidae_reduced_response <- columbidae_reduced_input_species$urban_tolerance
names(columbidae_reduced_response) <- columbidae_reduced_input_species$jetz_scientific_name

head(columbidae_reduced_response)
```

Is the a phylogenetic signal for urban tolerance in the columbidae family with species in at least 2 regional pools
```{r}
col_ut_reduced_phylo_signal <- phylosig(tree_pruned_to_columbidae_reduced, columbidae_reduced_response, method="lambda", test=TRUE)
col_ut_reduced_phylo_signal
```

Are species in low numbers of regional pool mostly present or absent?
```{r}
ggplot(col_ut[col_ut$region_count < 5,], aes(x = urban_tolerance)) + geom_bar()
```

## Is presence explained by number of regions?
```{r}
summary(lm(formula = urban_tolerance ~ region_count, data = col_ut))
```

What about for low numbers of regional pools?
```{r}
summary(lm(formula = urban_tolerance ~ region_count, data = col_ut[col_ut$region_count < 5,]))
```

What about for higher numbers of regional pools?
```{r}
summary(lm(formula = urban_tolerance ~ region_count, data = col_ut[col_ut$region_count > 5, ]))
```

## What about species in at least 5 regional pools?
```{r}
col_ut5 = col_ut[col_ut$region_count >= 5,]
```

```{r}
columbidae5_input_species <- col_ut5[col_ut5$jetz_scientific_name %in% tree$tip.label,]
tree_pruned_to_columbidae5 <- ladderize(drop.tip(tree, setdiff(tree$tip.label, col_ut5$jetz_scientific_name)))

columbidae5_response <- columbidae5_input_species$urban_tolerance
names(columbidae5_response) <- columbidae5_input_species$jetz_scientific_name

head(columbidae5_response)
```

Is the a phylogenetic signal for urban tolerance in the columbidae family
```{r}
col_ut5_phylo_signal <- phylosig(tree_pruned_to_columbidae5, columbidae5_response, method="lambda", test=TRUE)
col_ut5_phylo_signal
```

Show the results for species
```{r}
col5_continuous_trait_evolution <- contMap(tree_pruned_to_columbidae5, columbidae5_response)

colourCount <- length(col5_continuous_trait_evolution$cols)
col5_continuous_trait_evolution$cols[1:colourCount] <- rev(col5_continuous_trait_evolution$col[1:colourCount])
```

```{r}
cte_5 = plot(col5_continuous_trait_evolution, outline=FALSE, fsize=0.3, lwd=1, leg.txt="Cities/Regional Pools", type = "fan")
cte_5
```

### Store CTE plots
```{r}
par(mfrow=c(1,2)) 
plot(col_continuous_trait_evolution, outline=FALSE, fsize=0.3, lwd=1, leg.txt="Cities/Regional Pools", type = "fan")
text(0, 65, "All species")
plot(col5_continuous_trait_evolution, outline=FALSE, fsize=0.3, lwd=1, leg.txt="Cities/Regional Pools", type = "fan")
text(0, 60, "Min. 5 regional pools")
```
```{r}
jpeg("outputs/phylo.jpg", width = 2056, height = 1024)
par(mfrow=c(1,2)) 
plot(col_continuous_trait_evolution, outline=FALSE, fsize=0.8, lwd=3, leg.txt="Cities/Regional Pools", type = "fan")
text(0, 46, "All species", cex=2)
plot(col5_continuous_trait_evolution, outline=FALSE, fsize=1, lwd=3, leg.txt="Cities/Regional Pools", type = "fan")
text(0, 49, "Min. 5 regional pools", cex=2)
dev.off()
```

# Using Open Tree of Life
Load a maximum clade credibility tree from open tree
https://tree.opentreeoflife.org/opentree/opentree13.4@ott363030/Columbiformes

```{r}
library(rotl)
```

```{r}
colu <- read_csv('./bigquery_export__columbidae_birdlife_urban_tolerance.csv')
```


```{r}
taxa <- tnrs_match_names(unique(colu$species_name), context = "Birds")
```

```{r}
taxa[!is.na(taxa$unique_name),]
```

```{r}
opentree <- tol_induced_subtree(ott_id(taxa)[is_in_tree(ott_id(taxa))])
opentree
```

```{r}
plot(opentree, show.tip.label = FALSE)
```

```{r}
opentree$tip.label[1:4]
```

```{r}
taxon_map <- structure(taxa$search_string, names = taxa$unique_name)
taxon_map['Streptopelia decaocto']
```

```{r}
otl_tips <- strip_ott_ids(opentree$tip.label, remove_underscores = TRUE)
opentree$tip.label <- taxon_map[ otl_tips ]
opentree$node.label <- NULL
opentree$tip.label[1:4]
```


Check for species from data not in tree
```{r}
colu$species_name <- tolower(colu$species_name)
colu[!colu$species_name %in% opentree$tip.label,]
```

Format data
```{r}
columbidae_2_input_species <- colu[colu$species_name %in% opentree$tip.label,]
opentree_pruned_to_columbidae <- ladderize(drop.tip(opentree, setdiff(opentree$tip.label, colu$species_name)))

columbidae_2_response <- columbidae_2_input_species$urban_tolerance
names(columbidae_2_response) <- columbidae_2_input_species$species_name

head(columbidae_2_response)
```

There are no branch lengths in Open Tree of Life, so this does not work:
```{r eval=FALSE}
colu_phylo_signal <- phylosig(opentree_pruned_to_columbidae, columbidae_2_response, method="lambda", test=TRUE)
colu_phylo_signal
```

There are no branch lengths in Open Tree of Life, so this does not work:
```{r eval=FALSE}
col2_continuous_trait_evolution <- contMap(opentree_pruned_to_columbidae, columbidae_2_response)

colourCount <- length(col2_continuous_trait_evolution$cols)
col2_continuous_trait_evolution$cols[1:colourCount] <- rev(col2_continuous_trait_evolution$col[1:colourCount])
```
