---
title: "Trait predictors for presence"
output: html_notebook
---

```{r setup}
library(dplyr)
library(tidyverse)
library(ggpubr)
library(grid)
rm(list = ls())

col_ut <- read_csv('./bigquery_export__columbidae_jetz_urban_tolerance.csv')
mapping <- read_csv('./bigquery_export__columbidae_taxonomic_mapping.csv')
traits <- read_csv('./bigquery_export__columbidae_feeding_guilds.csv')

pigeons = left_join(col_ut, mapping)
pigeons = left_join(pigeons, traits, by = c('species_name' = 'scientific_name'))
pigeons = pigeons[!is.na(pigeons$trophic_niche),c('jetz_scientific_name', 'region_count', 'city_count', 'urban_tolerance', 'trophic_niche', 'foraging_niche', 'pc1', 'pc2', 'pc3', 'pc4')]
pigeons$trophic_niche = as.factor(pigeons$trophic_niche)
pigeons$foraging_niche = as.factor(pigeons$foraging_niche)

pigeons5 = pigeons[pigeons$region_count >= 5,]
```

```{r}
head(pigeons)
```
```{r}
nrow(pigeons)
```

```{r}
nrow(pigeons5)
```

```{r}
shapiro.test(pigeons$urban_tolerance) #p <= 0.05   -> Significantly different from normal
```

----------------------------
First look at feeding niches
----------------------------
```{r}
figure1 = ggarrange(ncol = 2, nrow = 1,
  ggplot(pigeons, aes(x = trophic_niche, y = urban_tolerance)) + 
    geom_boxplot() + labs(title = "All species") + rremove("ylab") + rremove("xlab"),
  ggplot(pigeons5, aes(x = trophic_niche, y = urban_tolerance)) + 
    geom_boxplot() + labs(title = "Min. 5 regional pools") + rremove("ylab") + rremove("xlab")
)

annotate_figure(figure1, left = textGrob("Urban tolerance", rot = 90, vjust = 1, gp = gpar(cex = 1.3)),
                    bottom = textGrob("Trophic niche"))
```

```{r}
kruskal.test(pigeons$urban_tolerance~pigeons$trophic_niche)
pairwise.wilcox.test(pigeons$urban_tolerance, pigeons$trophic_niche)
```

# Urban tolerance was significantly different between different trophic niches (Chi-squared=20.315, df=2, p<0.001), with urban tolerance significantly 
# lower in Frugivores compared to Granivores.

```{r}
kruskal.test(pigeons5$urban_tolerance~pigeons5$trophic_niche)
pairwise.wilcox.test(pigeons5$urban_tolerance, pigeons5$trophic_niche)
```
# Urban tolerance was significantly different between different trophic niches (Chi-squared=19.238, df=2, p<0.001), with urban tolerance significantly 
# lower in Frugivores compared to Granivores.

```{r}
figure2 = ggarrange(ncol = 2, nrow = 1,
  ggplot(pigeons, aes(x = foraging_niche, y = urban_tolerance)) + 
    geom_boxplot() + labs(title = "All species") + rremove("ylab") + rremove("xlab"),
  ggplot(pigeons5, aes(x = foraging_niche, y = urban_tolerance)) + 
    geom_boxplot() + labs(title = "Min. 5 regional pools") + rremove("ylab") + rremove("xlab")
)

annotate_figure(figure2, left = textGrob("Urban tolerance", rot = 90, vjust = 1, gp = gpar(cex = 1.3)),
                    bottom = textGrob("Foraging niche"))
```

```{r}
kruskal.test(pigeons$urban_tolerance~pigeons$foraging_niche)
pairwise.wilcox.test(pigeons$urban_tolerance, pigeons$foraging_niche)
```
# Urban tolerance was significantly different between different foraging niches (Chi-squared=24.204, df=3, p<0.001), with urban tolerance significantly 
# higher in ground feeding granivores.

```{r}
kruskal.test(pigeons5$urban_tolerance~pigeons5$foraging_niche)
pairwise.wilcox.test(pigeons5$urban_tolerance, pigeons5$foraging_niche)
```
# Urban tolerance was significantly different between different foraging niches (Chi-squared=19.238, df=2, p<0.001), with urban tolerance significantly 
# higher in ground feeding granivores.

Granivore outliers
```{r}
pigeons5[pigeons5$trophic_niche == 'Granivore' & pigeons5$urban_tolerance <= 0.25,]
```

Frugivore outliers
```{r}
pigeons5[pigeons5$trophic_niche == 'Frugivore' & pigeons5$urban_tolerance >= 0.75,]
```

-----------------------
What about morphology?
-----------------------
Check if PC traits are normal dist.
```{r}
shapiro.test(pigeons$pc1) #p <= 0.05   -> Significantly different from normal
```

```{r}
shapiro.test(pigeons$pc2) #p <= 0.05   -> Significantly different from normal
```

```{r}
shapiro.test(pigeons$pc3) #p > 0.05    -> NOT significantly different from normal
```

```{r}
shapiro.test(pigeons$pc4) #p > 0.05    -> NOT significantly different from normal
```


```{r}
figure3 = ggarrange(ncol = 2, nrow = 2,
  ggplot(pigeons, aes(x = pc1, y = urban_tolerance)) + 
    geom_point() + labs(title = "PC1") + rremove("ylab") + rremove("xlab"),
  ggplot(pigeons, aes(x = pc2, y = urban_tolerance)) + 
    geom_point() + labs(title = "PC2") + rremove("ylab") + rremove("xlab"),
  ggplot(pigeons, aes(x = pc3, y = urban_tolerance)) + 
    geom_point() + labs(title = "PC3") + rremove("ylab") + rremove("xlab"),
  ggplot(pigeons, aes(x = pc4, y = urban_tolerance)) + 
    geom_point() + labs(title = "PC4") + rremove("ylab") + rremove("xlab")
)

annotate_figure(figure3, left = textGrob("Urban tolerance", rot = 90, vjust = 1, gp = gpar(cex = 1.3)), top = textGrob("All species"))
```

```{r}
cor.test(pigeons$urban_tolerance, pigeons$pc1, method="spearman")
```
# There is a significant negative correlation between urban tolerance and PC1
# (Spearman's; n=144, p = 0.03, rho=-0.1807745)

```{r}
cor.test(pigeons$urban_tolerance, pigeons$pc2, method="spearman")
```
# There is NOT a significant correlation between urban tolerance and PC2
# (Spearman's; n=144, p = 0.393, rho=-0.0716447)

```{r}
cor.test(pigeons$urban_tolerance, pigeons$pc3, method="spearman")
```
# There is NOT a significant correlation between urban tolerance and PC3
# (Spearman's; n=144, p = 0.67, rho=0.03602794)

```{r}
cor.test(pigeons$urban_tolerance, pigeons$pc4, method="spearman")
```

# There is NOT a significant correlation between urban tolerance and PC4
# (Spearman's; n=144, p = 0.17, rho=-0.115925)

```{r}
figure4 = ggarrange(ncol = 2, nrow = 2,
  ggplot(pigeons5, aes(x = pc1, y = urban_tolerance)) + 
    geom_point() + labs(title = "PC1") + rremove("ylab") + rremove("xlab"),
  ggplot(pigeons5, aes(x = pc2, y = urban_tolerance)) + 
    geom_point() + labs(title = "PC2") + rremove("ylab") + rremove("xlab"),
  ggplot(pigeons5, aes(x = pc3, y = urban_tolerance)) + 
    geom_point() + labs(title = "PC3") + rremove("ylab") + rremove("xlab"),
  ggplot(pigeons5, aes(x = pc4, y = urban_tolerance)) + 
    geom_point() + labs(title = "PC4") + rremove("ylab") + rremove("xlab")
)

annotate_figure(figure4, left = textGrob("Urban tolerance", rot = 90, vjust = 1, gp = gpar(cex = 1.3)), top = textGrob("Min. 5 regional pools"))
```

```{r}
cor.test(pigeons5$urban_tolerance, pigeons5$pc1, method="spearman")
```

# There is a significant negative correlation between urban tolerance and PC1
# (Spearman's; n=114, p = 0.015, rho=-0.2272254)

```{r}
cor.test(pigeons5$urban_tolerance, pigeons5$pc2, method="spearman")
```

# There is NOT a significant correlation between urban tolerance and PC2
# (Spearman's; n=114, p = 0.15, rho=-0.136747)

```{r}
cor.test(pigeons5$urban_tolerance, pigeons5$pc3, method="spearman")
```

# There is NOT a significant correlation between urban tolerance and PC3
# (Spearman's; n=114, p = 0.57, rho=0.05381857)

```{r}
cor.test(pigeons5$urban_tolerance, pigeons5$pc4, method="spearman")
```

# There is NOT a significant correlation between urban tolerance and PC4
# (Spearman's; n=114, p = 0.36, rho=-0.0857774)

--------------------------------------
so we have a model like:
--------------------------------------
```{r}
model <- glm(formula = urban_tolerance ~ pc1 + trophic_niche, data = pigeons)
summary(model)
```

```{r}
model_reduced <- glm(formula = urban_tolerance ~ trophic_niche, data = pigeons)
summary(model_reduced)
```

R-squared:
```{r}
with(summary(model_reduced), 1 - deviance/null.deviance)
```

And for species in 5 regional pools or more...
```{r}
model5 <- glm(formula = urban_tolerance ~ pc1 + trophic_niche, data = pigeons5)
summary(model5)
```

```{r}
model5_reduced <- glm(formula = urban_tolerance ~ trophic_niche, data = pigeons5)
summary(model5_reduced)
```

R-squared:
```{r}
with(summary(model5_reduced), 1 - deviance/null.deviance)
```
